---
title: "Verify Wind"
author: "REsurety, Inc."
date: "February 2016"
output: html_document
---


## REsurety Software Development Task

We have received a new cluster of wind data sources that are near Logan Wind Energy, an operating wind
project.  Your goal is to analyze whether the new wind data can be used
to model long-term conditions at the wind farm. Your answer should have a quantitative
statistical answer and a visual component.

Your data sources are:

* `WindData.csv` - hourly wind speed, direction, and temperature back to 1979
* `SiteData.csv` - data about Logan such as owner, location, and turbine specifications
* `PowerCurve.csv` - relates wind speed to turbine power production and thrust
* `SiteEnergy.csv` - shows historical energy production at Logan

The following sections describe each data source in detail and describe the 
expected deliverables. Expected time for the task is ~2 hours.

## Data Source #1: Wind Time Series

The file `WindData.csv` contains wind data from a meteorological site at latitude 41.0 and longitude -103.3333,
to the North and West of Logan.  Columns in the file are:

* `Time` - hourly data going back to 1979 in UTC
* `NW_spd_w2`, `NW_spd_w10`, `NW_spd_50m` Wind speed at 2m, 10m, and 50m (m/s)
* `NW_dir_w2`, `NW_dir_w10`, `NW_dir_50m` Wind direction at 2m, 10m, and 50m (degrees, 0 = North)
* `NW_t2m` - Temperature at 2m


```{r, echo=FALSE, results='hide', fig.show='hold'}
# For the convenience of R users, we have also provided `WindData.rds` which has the same information
# but is much faster to load.
wind <- data.table::data.table(readRDS(file='WindData.rds'))
wDen10 <- density(wind$NW_spd_10m,cut = 0)
wDen02 <- density(wind$NW_spd_2m,cut = 0)
wDen50 <- density(wind$NW_spd_50m,cut = 0)
par(mar=c(4,4,4,4),mgp=c(2,0.5,0))
plot(wDen10$x,wDen10$y,type='l',lwd=2,col="#0072B5",xlab='Wind Speed (m/s)',
     ylab='Frequency',yaxt='n',main='Distribution of Wind by Height')
lines(wDen50$x,wDen50$y,lwd=2,col='#84BE41')
lines(wDen02$x,wDen02$y,lwd=2,col='#808080')
legend('topright',legend=c('50m','10m','2m'),lty='solid',lwd='2',col=c('#84BE41','#0072B5','#808080'))
```

## Data Source #2: Site Characteristics

The wind project we are evaluating has all its site characteristics
saved in the file `SiteData.csv`:

```{r, echo=FALSE}
read.csv(file='SiteData.csv',stringsAsFactors=F,row.names=1)
```


## Data Source #3: Turbine Power Curve

A wind turbine's power curve shows the idealized relationship between incoming 
wind and the amount of power produced.  The file `PowerCurve.csv` contains columns
 `WindSpeed.ms`, `Power.kW`, and `ThrustCoefficient` and is valid for the site's
 GE 1.5SLE turbine.  Like most power curves, the turbine produces no power below
 its ~3.5m/s cut-in speed then rises almost linearly to its nameplate capacity
 of 1500 kW.  Beyond 25m/s the turbine hits its cut-out speed and produces no power.

```{r, echo=FALSE, results='hide', fig.show='hold'}
pwr  <- read.csv(file='PowerCurve.csv')
par(mar=c(4,4,4,4))
plot(pwr$WindSpeed.ms,pwr$Power.kW,type='l',col="#0072B5",lwd=2,xlab='Wind Speed (m/s)',ylab='Power (kW)',main='GE 1.5SLE Power Curve')
```


## Data Source #4: Site Energy Production

The file `SiteEnergy.csv` contains public records for Logan Wind Energy's monthly
energy production.  Columns are:

* `Month` - YYYY-MM of record
* `Energy.MWh` - Megawatt-hours of reported site production in that month
* `Capacity.MW` - The total nameplate capacity (in MW) reported by the site

```{r, echo=FALSE, fig.show='hold'}
eObs <- read.csv(file='SiteEnergy.csv',stringsAsFactors=F)
eObs$Year <- as.numeric(substr(eObs$Month,1,4))+(as.numeric(substr(eObs$Month,6,7))-1)/12
par(mar=c(4,4,4,4))
plot(eObs$Year,eObs$Energy.MWh,type='l',col="#0072B5",lwd=2,xlab='Year',
     ylab='Monthly Energy (MWh)',main='Logan Wind Energy Site Production')
tail(eObs)
```


## Hints

Your workflow may resemble the following:

* Transform an appropriate hourly wind speed using the power curve to get hourly energy production
* Aggregate hourly energy production to the monthly level
* Compare modeled monthly energy to the publicly reported observed energy

## Deliverables

Submit the following to `kcritz@resurety.com` as PDF or HTML files:

* Your contact information
* A 3-4 sentence answer to the question "Is the met data suitable to
  predict energy output at the site?" This answer should have a quantitative
  component.
* A visualization that illustrates your point.  Charts will be evaluated
  according to Tufte's "Visual Display of Quantitative Information"
  <http://www.edwardtufte.com/tufte/books_vdqi>
* The code you used to answer the question. Use any analysis environment that 
  works for you. If you use Excel, we will laugh at you.
* Is there anything we can do to improve this exercise?  Was anything unclear
  or ambiguous?

Expectations:

* Most candidates spend no more than 2 hours on this task.  This isn't meant to be
  an in-depth consulting project or second job.  Obviously we can't enforce
  a time limit, but you shouldn't let this eat your time.
* Results should reflect your work only. This is not a group project.  
* If you have domain knowledge, don't worry about nonlinear shear, time
  zone corrections, or wake/performance losses.  This investigation is meant
  to be approximate and fast. 
  No extra credit will be given if you build a terrain map and locate 
  turbines to run a full CFD simulation of the site.
  
  









